---
title: "Explore gapminder & use dplyr"
author: "Lendi N. Joy"
date: "2/9/2021"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load Necessary Libraries 
```{r warning=FALSE}
library(ggplot2)
library(gapminder)
library(dplyr)
```

###Exercise 1: Basic dplyr

## 1.1
Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.
```{r}
filter(gapminder, country=="Brazil" | country=="China" | country=="Niger", year>=1970&year<=1979)
```

## 1.2
Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.
```{r}
gapminder %>% 
  select(country, gdpPercap)
```

## 1.3
Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.

Hint: you might find the lag() or diff() functions useful.
```{r}
gapminder %>% 
  mutate(lifeExp_change = lifeExp-lag(lifeExp))%>% 
  filter(lifeExp_change<0)
```

## 1.4
Filter gapminder so that it shows the max GDP per capita experienced by each country.
_Hint: you might find the max() function useful here._
```{r}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap== max(gdpPercap))
```

## 1.5
Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.
```{r}
gapminder %>% 
  filter(country == "Canada") %>%  
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point()+
  scale_y_log10(labels = scales::comma_format())
```

#Exercise 2: Explore two variables with dplyr and ggplot2
Use palmerpenguins::penguins or another dataset of your choice. _(Check out a dataset from the datasets R package if you want!)_

##2.1
Pick two quantitative variables to explore.
1. Make a summary table of descriptive statistics for these variables using summarize().
  * Include whatever staistics you feel appropriate (mean, median sd, range, etc.).
2. Make a scatterplot of these variables using ggplot().
```{r}
palmerpenguins::penguins%>% 
  summarize(Flipper_mean=mean(flipper_length_mm, na.rm=TRUE), 
            flipper_sd=sd(flipper_length_mm, na.rm=TRUE),
            mass_mean=mean(body_mass_g, na.rm=TRUE),
            mas_sd=sd(body_mass_g, na.rm=TRUE))
ggplot(palmerpenguins::penguins, aes(body_mass_g, flipper_length_mm))+
  geom_point()
```

## 2.2
Pick one categorical variable and one quantitative variable to explore.
1. Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group.
2. Make one or more useful plots to visualize these variables.
```{r}
palmerpenguins::penguins%>% 
  group_by(species, island, year) %>% 
  summarize(flipper_mean=mean(flipper_length_mm, na.rm=TRUE), 
            flipper_sd=sd(flipper_length_mm, na.rm=TRUE),
            mass_mean=mean(body_mass_g, na.rm=TRUE),
            mas_sd=sd(body_mass_g, na.rm=TRUE), n()) %>% 
  ggplot(aes(mass_mean, flipper_mean))+
  geom_point() %>% 
```

#Bonus Exercise: Recycling (Optional)
Evaluate this code and describe the result. The goal was to get the data for Rwanda and Afghanistan. Does this work? Why or why not? If not, what is the correct way to do this?

filter(gapminder, country == c("Rwanda", "Afghanistan"))

