# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

Load the data.
```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
```

Load necessary libraries.
```{r}
library(dplyr)
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
```{r}
summary(green_data)
```
Several variables appear to have missing values coded as -99, so we will remove those.

```{r}
green_data <- na_if(green_data, "-99") 
summary(green_data)
```

  2. Compute total scores for the four scales. 
     Recode variables as needed.
```{r}
reverse <- function(x) (6 - x)
green_data_comp <- green_data %>% 
   rowwise() %>% 
  mutate(across(c(green2, comp6:comp10, intel7:intel10, open7:open10), reverse)) %>% 
  mutate(green_mean = mean(c_across(starts_with("g")), na.rm = TRUE), comp_mean = mean(c_across(starts_with("c")), na.rm = TRUE), intel_mean = mean(c_across(intel1:intel10), na.rm = TRUE), open_mean = mean(c_across(starts_with("o")), na.rm = TRUE)) %>% 
  ungroup()
```

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
```{r}

```
     
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
