---
title: "Final Project"
author: "Lendi Nicole Joy"
date: "5/3/2021"
output: 
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.width = 10,
                      fig.align = "center",
                      fig.path = here::here("output", 
                                            "figures"))
```

### Load the needed libraries.  
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(psych)
library(GPArotation)
library(flextable)
```

### Load the dataset.   
```{r, message = FALSE}
Motivation <- read_csv(here::here("data", "Activities_and_Mood.csv"))

```

### Clean the data.  
Next, clean up the data using variable Q166_7, which indicates whether the survey was completed.
```{r}
Motivation <- 
  Motivation %>% 
  drop_na("Q166_7")
```
There were 149 cases removed.

Now let's select only the variables needed for the EFA (remove V4 and V5, which are attention check items).
```{r}
Motivation_EFA <- 
  Motivation %>% 
  select(VFI.U1:Int_4, 
         -c(V4, V5))  
```

##### Reverse code items.  
After selecting the appropriate variables, I will reverse code the items so that higher scores indicate more positive attitudes. A few items are reverse coded, so exclude those variables from recoding.
```{r}
Motivation_EFA<-  
  8 - Motivation_EFA[, -c(31, 33:35, 52, 57)]
```

### Save the dataset.  
Save the resulting dataset.
```{r}
write.csv(Motivation_EFA,
          here::here("data", "Motivation_EFA.csv"))
```

### Run the EFA.   
The original hypothesis is that there will be 7 belief factors, 3 factors indicating attitudes, subjective norms and perceived behavior control, then 1 factor to indicate intention or motivation to volunteer.

I start with extracting these 11 factors since this is the original hypothesis.
```{r}
EFAresult1 = factanal(~ ., 
                      data = Motivation_EFA, 
                      factors = 11, 
                      fm = "pa", 
                      rotation = "oblimin")
EFAresult1
```
##### Determine number of factors.   
We look at the sum of squared loadings (SS loadings), which are the eigenvalues. The Kaiser Rule suggests that eigenvalues greater than 1 indicate meaningful factors. This rule, would suggest that there are *10 meaningful* factors.


We can look at the item loadings and see that even though ten factors have eigenvalues greater than 1, the items do not load very highly on some of the factors.

###### Use another method to determine number of factors.  
Next, I will run a parallel analysis and examine the scree plot as another way of determining the number of factors to extract.
```{r Parallel Analysis, fig.show = "hide"}
set.seed(13115)
PA = fa.parallel(Motivation_EFA, 
                 fa = "fa",
                 quant = .95)
```

Instead of displaying the scree plot that is created, I will make a nicer plot of the parallel analysis.

First, create a data frame with the observed values.
```{r}
obs = data.frame(PA$fa.values)
obs$type = c('Observed Data')
obs$num = c(row.names(obs))
obs$num = as.numeric(obs$num)
colnames(obs) = 
  c('eigenvalue', 
    'type', 
    'num')
```

Next, calculate and save the 95% quantiles for the eigenvalues for the observed data.
```{r}
percent = apply(PA$values,
                2,
                function(x) 
                  quantile(x,.95))
min = as.numeric(nrow(obs))
min = (4*min) - (min-1)
max = as.numeric(nrow(obs))
max = 4*max
percentile = percent[min:max]
```

Now create a data frame for the simulated eigenvalues.
```{r}
sim = data.frame(percentile)
sim$type = c('Simulated Data (95th %ile)')
sim$num = c(row.names(obs))
sim$num = as.numeric(sim$num)
colnames(sim) = 
  c('eigenvalue', 
    'type', 
    'num')
```

After creating the appropriate data frames, I will put them together into one data frame.
```{r}
PAdata = rbind(obs,sim)
```

Create the plot.
```{r Pretty Scree Plot}

ggplot(PAdata, 
                  aes(x = num, 
                      y = eigenvalue, 
                      shape = type)) +
geom_line()+
geom_point(size = 1.5)+
scale_y_continuous(name = 'Eigenvalue')+
scale_x_continuous(name = 'Factor Number', 
                   breaks = min(PAdata$num):max(PAdata$num))+
scale_shape_manual(values = 
                     c(16,5)) +
geom_vline(xintercept = PA$nfact, linetype = 'dashed') +
  ggtitle("Scree Plot")
```

### Rerun EFA.  
Now I will rerun the EFA with only 8 factors using the "fa" function from the psych package, but I will not print the results just yet.
```{r}
EFAresult2 = fa(Motivation_EFA, 
                      nfactors = 8, 
                      fm = "pa", 
                      rotate = "oblimin")
```

Use the results from the EFA to create a table of the factor loadings.
```{r message=FALSE}
table <- function(x, cut) 
  {loadings <- fa.sort(x)$loadings %>% round(3)
  loadings[loadings < cut] <- ""
  tableinfo <- cbind(x$communalities, 
                    x$complexity) %>%
    as.data.frame() %>%
    rename("Communality" = V1,
           "Complexity" = V2) %>%
        rownames_to_column("Item")
  loadings %>%
    unclass() %>%
    as.data.frame() %>%
    rownames_to_column("Item") %>%
    left_join(tableinfo) %>%
    mutate(across(where(is.numeric), round, 3))
  }  
   
fa_table <- flextable(table(EFAresult2, .30))
```

```{r echo=FALSE}
fa_table %>% 
set_caption(
  caption = "EFA Factor Loadings",
  autonum = TRUE,
  style = "Table Caption",
  html_escape = TRUE) %>% 
  autofit()
```

```{r include=FALSE}
save_as_image(x = fa_table, path = here::here("output", "EFA.png"), webshot = "webshot")
```