---
title: "HW01 Exploring Data"
author: "Lendi N. Joy"
date: "1/25/2021"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load and Examine
First we need to load the psych package and the data. Let's also look at the structure of the data at this time.
```{r warning=FALSE}
library(psych)
data(bfi)
str(bfi)
```
Alright, 2,800 cases and 28 variables. 

### Add Value Labels
We saw that gender and education are coded as integer variables, so let's add some value labels.  
**gender**:    1 = male, 2 = female   
**education**: 1 = HS, 2 = finished HS, 3 = some college, 4 = college graduate 5 = graduate degree
```{r}
bfi$gender<- factor(bfi$gender, levels = c(1,2), labels = c("male", "female"))
bfi$education<- factor(bfi$education, levels = c(1,2,3,4,5), labels = c("HS", "finished HS", "some college", "college grad", "grad degree"))
```

### Check Missing Data
We also saw that there were some missing values when we looked at the structure. Let's find out how many complete cases there are in the dataset.
```{r}
sum(complete.cases(bfi)) #Complete cases
sum(!complete.cases(bfi)) #Incomplete cases
```

### Clean it up
It's probably not good practice to just throw out data, but let's do it anyway and keep only complete cases.
```{r}
x <- na.omit(bfi)
```


Interestingly, there are some participants under 18.   
Let's find out how many now that we are only including complete cases.
```{r}
sum(x$age < 18)
```
OK, only 51. Let's pretend that we have parental consent and keep them in.

### Summarize
Let's get a summary of our complete data.
```{r}
summary(x)
```

Lastly, let's take a look at education level by gender.
```{r}
table(x$gender, x$education)
```